Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-05 manuscript-chicago.Rmd
Snipping words

abstract: |
  In recent years, individuals involved in social movements have taken to social
  media platforms to reach broader audiences at rapid speeds. Although
  micro-blogging has become a primary means of information dissemination during
  times of turmoil, the degree to which virtual mobilization reflects or drives
  real-world movements is unclear. Here, we explore the dynamics of real-world
  events and Twitter social cohesion in Syria during the Arab Spring. Using the
  nonlinear methods cross-recurrence quantification analysis and windowed
  cross-recurrence quantification analysis, we investigate if frequency of
  events of different intensities are coupled with social cohesion found in
  Syrian tweets. Results indicate that virtual social cohesion is coupled with
  the counts of all and negative events each day but shows a more nuanced
  connection to positive events, suggesting that virtual mobilization does
  reflect real-world movements. We conclude with a discussion of implications
  and applications of nonlinear methods in political science research.
  <!-- https://tinyurl.com/ybremelq -->

@@

Prior to the advent of social media, people coordinated their protest behavior
using phone tree calling sheets and mimeographed flyers, through in-person
meetings of social organizations and churches, and through word of mouth. They
gathered together and hand-made protest signs that reflected their agendas. As
McAdam found, people were more likely to join movements when they had a personal
connection [@mcadam1986recruitment; @mcadam1993specifying]: Historically, strong
ties have been strong predictors of involvement in high-risk protest activity.

The cohesiveness and strength of social movements in the pre-Internet era could
be operationalized by the number of people turning out to protest and through
the political events that precipitated and resulted from social gatherings *en
masse*. Cohesive social movements were reflected in their cohesive placards and
posters. For example, the “I am a man” posters held by the men involved in the
1968 sanitation workers' strike reflected their internal cohesion and unified,
goal-oriented movement [@Windsor2020].

However, the rise of the Internet and its decentralized online networks has
upended theories about recruitment to---and participation in---social movements.
Strong interpersonal ties have yielded to weak impersonal ties that strongly
predict involvement in social movements [@steinert2017spontaneous]. Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to “big data” analyses of language on social media.

Today, social media replicates and amplifies this phenomenon during moments of
social unrest. Amid the cacophony of messages during political upheaval---such
as the Arab Spring---consistent themes emerge through hashtags, retweets, and
the repeated mentions of concepts such as regime, demonstration, and
human rights. The cohesiveness of these themes (i.e., whether movement
participants are metaphorically singing in one voice) reflects the real-world
events happening contemporaneously with collective action in social movements or
civil conflicts.

@@

Disentangling the complicated relationship between mass online
mobilization---and social mobilization in person---continues to be relevant,
even as our modes of communication change. The political events that
transpired during and after the Arab Spring are still in the process of being
understood. As Zeitzoff [-@zeitzoff2017social] recognizes, social media is
changing conflict, but the mechanisms and relationships by which it does so are
still largely unknown.

## Syria in Context
The decades preceding the Arab Spring in Syria were marked by
social repression and fear. In 1971, Hafez al-Assad overthrew the previous
leader and remained in power for thirty years, presiding during
regional contentiousness and cooperation until his death in 2000. His son,
Bashar al-Assad continued the regime of authoritarian leadership as he assumed
his office. The Assad lineage hails from a minority Islamic sect (the Alawites)
in a multi-religious and multi-ethnic country comprising Christians, Sunni and
Shia Muslims, Kurds, and Palestinian refugees.

@@

Following a wave of democratization efforts across the Middle East and North
Africa, modest Syrian protests began in March 2011. The regime responded harshly
with imprisonments, detentions, censorship, and military operations as the
resistance to Assad's regime grew and spread throughout the country. The
fighting escalated dramatically in March 2012, after increased unrest in the
western city of Homs. International actors---including the Red Cross/Red
Crescent Society, the United Nations General Assembly, the Security Council, and
individual world leaders---worked toward cease-fires, negotiations, and
concessions without much success. While many high-level military and diplomatic
officials defected from the regime, numerous loyal forces continued the regime's
repression, and the Syrian people turned to social media to help coordinate
their activism and communicate to the world. Between March and June 2012, more
than 8 million tweets were sent within Syria, revealing the details of the
conflict suppressed by the government.

Syria is a multi-ethnic, multi-lingual, and multi-religious state. Kurdish
people make up between 10-15% of the population, Alawites between 8-15%,
Christians about 10%. There are also much smaller sects (e.g., Ismailis, Druze),
but Sunni Muslims comprise the vast majority of society. Most people speak
Arabic as their primary language. Most policy reports addressing the current
conflict report on the cohesion of the ruling Alawite military elite, rather
than on the cohesiveness of the opposition movement. However, the Sunni
majority's disaffection with the Syrian government and their lack of political
representation provide sufficient grievance and collective group identification
to make them a cohesive force.

@@

Massive online social cohesion can have real-world social manifestations. Social
media platforms provide one way to enable this kind of online social cohesion by
offering distributed information-gathering and real-time information
dissemination. Given its accessibility and proficiency as an information
exchange tool, Twitter has provided individuals involved in social movements an
opportunity to complement or even replace real-world social mobilization. For
example, during the 2012 Italian protests for global economic change, the riots
in Rome were mirrored by Twitter activity both on the day of the protests and in
the days following them [@vicari2013public]. Tweeting was determined to be more
effective locally in discussing real-world, real-time events than traditional
mainstream media platforms. As more and more individuals rely on social media to
obtain local news and information, it becomes increasingly important to
understand just how these micro-blogs reflect real-world action. Some argue that
these mass movements would not have been possible without efficient large-scale
coordination, both physically on the ground and virtually on social media
platforms.

Along with Facebook, Twitter has been credited as essential to the Arab
Spring---both for protesters (as a means to coordinate with one another and to
communicate with the world) and for the government response (to monitor and
track protestors to crack down on dissent) [@comninos2011twitter]. The events of
Arab Spring have been thoroughly discussed within the political science
literature [e.g., @lutterbeck2013arab;
@koehler2017political], but to our knowledge, the direct *coupling* between
Syrian Twitter cohesion and international action coordination has not been
investigated. In fact, some have been skeptical about the effect of social media
on collective action at all [@seay2014slacktivism].

This skepticism can be summarized in a quote by Lisa @anderson2011demystifying:

@@

Interestingly, these views may
actually help to make the case for bridging the gap between social media and
social mobilization. Such views suggest the strength of social ties in an online
environment is inextricably linked to the *context* and *ambitions* of the actors.
Moreover, research suggests that weak ties still retain considerable political
power: While strong ties tend to bind smaller groups, weak ties
tend to bind groups across distances [@granovetter1977strength].

Social media generates critical information about events, including whether and
how a demonstration will take place, how many people participate, and how those
in power (e.g., state, military, security apparatus) respond. Similarly, these
media platforms alert us to more global shifts within a conflict, such as when
fighting moves to a new city or region, the number of casualties and wounded,
and the types of weaponry used by parties [@zeitzoff2017social]. In monitoring
social media channels, government forces may respond to the perception that
citizens are gaining strength and sympathy. Governments have been shown to be
quite tolerant of dissent so long as it does not metamorphose into collective
action [@king2013censorship].

Citizens communicate with each other with increased transparency due to the
ubiquitous availability of social media such as low-cost and egalitarian
platforms like Twitter. Citizens are able to access these tools via mobile
phones and computers with Internet connectivity, providing a stream of real-time
shared information. Citizens can share the location of events like protests and
military incursions, the degree of state repression, and the number of
casualties. In turn, the government---interested in remaining in
power---monitors social media platforms to ascertain emerging threats
[@comninos2011twitter], such as where new groups are forming, where existing
groups are strengthening, where gatherings are held, and who the idea
entrepreneurs are.

@@

\setlength{\leftskip}{0pt}
\end{singlespace}
\noindent With few exceptions, Assad chose to keep social media platforms up and running.
Gohdes [-@gohdes2015pulling] notes that regimes can attempt to thwart the growth
of social protest online and in the streets by shutting down the Internet.
During these times, however, they find that military repression increases.
However, government information about the citizens decreases; as such, many
regimes have elected to maintain Internet connectivity as a means of monitoring
and surveilling their citizens.

@@

The type of collective contentious behavior that manifests in society is a
function of the available political opportunity structures. The political
opportunity structure are “comprised of *(sic)* specific configurations of
resources, institutional arrangements and historical precedents for social
mobilization, which facilitate the development of protest movements in some
instances and constrain them in others” [@kitschelt1986political]. What emerges
as strategies and tactics in contentious politics are, in essence, a function of
what the system makes possible. This includes the number of people and their
disposition (i.e. nonviolent or violent), as well as their internal structure.
The more cohesive a group is, the greater threat they pose to the establishment.

Cohesion is also a familiar concept in the study of civil conflicts, as it
pertains to group affiliation and the ability to overcome collective action
problems. Individuals must act in concert---that is, cohesively---to
accomplish the task of challenging the authority of the state [@olson1965logic;
@eisinger1973conditions; @tilly2004contentious; @francisco1995relationship].
Recent work has considered rebel recruitment for participation in
civil conflicts as a function of internal cohesion [@gates2002recruitment]:
Rebel movement success is influenced by the degree of group cohesion, which
itself is a function of geographic distance, ideology, and ethnicity.
Within Syria, multiple rebel groups exist, sourcing fighters locally
from the disaffected Syrian population and internationally through
fighters from transnational terrorist organizations. There is considerable
intergroup competition for participation in the rebel movement and
punishment for defectors for nonparticipation [@gates2002recruitment].
Foreign fighters would appear to be amply motivated to fight, given their
decision to relocate from a home country that may be located in the region
or much farther away.

@@

The civil war literature has further explored the role of social cohesion in
detail. Social homogeneity increases the likelihood of social conflict,
especially when the society is polarized [@collier2004greed]; this is partly
due to nationalist groups' strong identification with a collective identity
[@cederman2007beyond] and their social solidarity [@sambanis2009s]. These
same processes can be similarly important for rebels [@anyanwu2002economic].
Referencing Gurr's [-@gurr2000peoples] work on minorities at
risk, Regan and Norton [-@regan2005greed] note that

@@

We contextualize linguistic cohesion with social mobilization and response to
the relative intensity of the conflict. Cohesion is a multi-layered concept
referring to the connectedness of speakers, actions, and societies. It
reflects the continuity of language and expression, and shows the
connectivity of participants in a social movement. To address the relationship
between language and discourse and social movements, we adopt an
interdisciplinary approach that incorporates linguistics, psychology, and
political science using nonlinear dynamical systems methodology to disentangle the
relationship between language and action.

@@

The words that comprise social media posts are the first level of cohesion that
we will consider.  In Figure \@ref(fig:cohesion-in-tweets), we see four
sequential tweets, each containing four words. Each sequential tweet contains
some content included in the previous tweets. The linguistic construct of
content word overlap measures the relationship across units of meaning---here,
tweets---where identical and semantically similar words are repeated
[@graesser2004coh]. These repeated lexical items constitute a theme; on Twitter,
this might be a particular hashtag, a name, a location, or other nominal
category. This localized cohesion is referential; words are repeated and
co-referred across tweets, creating a sense of continuity and cohesiveness of
meaning and message.

@@

In estimating cohesion in Twitter posts, tweets that share content words
will yield higher cohesion values, even among speakers with opposing
views on the subject matter. Repetition of terms and concepts is captured by
reiteration, which is a type of lexical cohesion in which a subsequent item
repeats a previous item, or refers to it through synonymous language
[@halliday1976cohesion]. In social media, we can conceive of retweets (i.e., reposts
or forwards of previous messages) as a form of reiteration
that increase message cohesion. Since retweets replicate verbatim some part of
another tweet, these tweets would also have higher cohesion. Hashtags function
similarly and are included in our analysis.

@@

The use of similar language can provide insight into *interlocutor cohesion*, or
the degree to which people are talking about similar things. This can include
people who are talking about similar things from radically perspectives, as well
as those who are talking about something from the same perspective. Previous
research from political science has investigated these ideas [@Windsor2020]: The
“Occupy Wall Street” movement has been held up as an example of a movement with
low interlocutor and linguistic cohesion because of its wide range of signs and
slogans, while the “I am a man” protest (captured by Ernest Withers during the
1968 sanitation workers' strike in Memphis, TN) or the “Black Lives Matter”
placards (following the uprising in Ferguson, MO) have high lexical and
interlocutor cohesion [@edgar2018struggle].

Fed by lexical cohesion, interlocutor cohesion ebbs and flows along a timeline
contextualized by variation in real world events. Interlocutor cohesion may
increase preceding an in-person social mobilization event, alerting a repressive
government that collective mobilization is happening. Interlocutor cohesion may
also respond to exogenous events such as spontaneous, unprovoked militarized
conflict.

Given the importance of lexical and interlocutor cohesion for messaging in
political movements, lexical cohesion may be important trace data for
researchers interested in leveraging social media data to analyze social
movements. Similar to large-scale studies of sociocultural movements [sometimes
referred to as *culturomics*; e.g., @michel2011quantitative], political
scientists can track the ways in which conversations on social media coalesce or
diffuse around similar content words. In this way, it matters relatively little
whether the interlocutors have precisely the same views: The point is to
understand whether the society-level conversation is converging on similar
language.

@@

At the same time, behavioral similarity in certain contexts has been
connected to important psychological and social effects. For example, shared
language is associated with rapport [both in face-to-face and computer-mediated
conversations; @niederhoffer2002linguistic; @riordan2011evidence] and
can even be connected to better performance on shared tasks [as long as
the language is relevant to the task at hand; @fusaroli2012coming]. Shared
movement is associated with greater group bonding and greater willingness
to contribute to the group [@wiltermuth2009synchrony], and shared identity
is associated with a greater willingness to adopt the group's goals and
motivations [@walton2012mere].

We should expect to see increased linguistic cohesion among people participating
in online discussions about an idea, event, individual, or movement. Given that
Syrian protesters are known to have coordinated their activity through Twitter
by discussing their own events (e.g., planned protests, active demonstrations)
and their reactions to events happening around them (e.g., government reactions,
international action or inaction), the cohesion in social media should be
expected to correlate with the social cohesion of the group [@zeitzoff2011using].

This social cohesion can have real and wide-ranging impacts on a political
movement and the people in it. A socially cohesive group is inherently more
threatening to the existing regime because it reflects collective opposition
momentum and the professionalization of potential alternate leadership. Global
insurgency movements exhibit similar patterns in their social structure
[@bohorquez2009common]. Rather than frenetic, patchwork messages via Twitter,
emergent or intentional cohesive online messaging---using similar language,
paralleling in-person messaging [@Windsor2020]---are a sign of increased
organizational structure, inciting a government response.

>>>>

abstract: |
  In recent years, political activists have taken to social media platforms to
  rapidly reach broad audiences. Despite the prevalence of micro-blogging in
  these sociopolitical movements, the degree to which virtual mobilization
  reflects or drives real-world movements is unclear. Here, we explore the
  dynamics of real-world events and Twitter social cohesion in Syria during the
  Arab Spring. Using the nonlinear methods cross-recurrence quantification
  analysis and windowed cross-recurrence quantification analysis, we investigate
  if frequency of events of different intensities are coupled with social
  cohesion found in Syrian tweets. Results indicate that virtual social cohesion
  is coupled with the counts of all and negative events each day but shows a
  more nuanced connection to positive events, suggesting that virtual
  mobilization does reflect real-world movements. We conclude with a discussion
  of implications and applications of nonlinear methods in political science
  research.
  <!-- https://tinyurl.com/ybremelq -->

@@

Before social media, people coordinated protests using phone tree calling sheets
and mimeographed flyers, through in-person meetings of social organizations and
churches, and by word of mouth. They gathered together and hand-made protest
signs that reflected their agendas. As McAdam found, people were more likely to
join movements when they had a personal connection [@mcadam1986recruitment;
@mcadam1993specifying]: Historically, strong ties have been strong predictors of
involvement in high-risk protest activity.

The cohesiveness and strength of social movements before the Internet were
operationalized by the number of people present and through the political events
that precipitated and resulted from social gatherings *en masse*. Cohesive
social movements were reflected in their cohesive placards and posters. For
example, the pervasive “I am a man” posters during the 1968 sanitation workers'
strike reflected their internal cohesion and unified, goal-oriented movement
[@Windsor2020].

However, the rise of the Internet and its decentralized online networks have
upended theories about recruitment to---and participation in---social movements.
Strong interpersonal ties have yielded to weak impersonal ties that strongly
predict involvement in social movements [@steinert2017spontaneous]. Measurements
of social movements have similarly evolved from ethnographic interviews to “big
data” analyses of social media.

Amid the cacophony of messages during political upheaval---such
as the Arab Spring---consistent themes emerge through hashtags, retweets, and
the repeated mentions of concepts such as regime, demonstration, and
human rights. The cohesiveness of these themes (i.e., whether movement
participants are metaphorically singing in one voice) reflects the real-world
events happening contemporaneously with collective action in social movements or
civil conflicts.

@@

Disentangling the complicated relationship between mass online
mobilization---and social mobilization in person---continues to be relevant,
even as our modes of communication change. Accordingly, the political events
around the Arab Spring are still being understood. As Zeitzoff
[-@zeitzoff2017social] recognizes, social media is changing conflict, but the
mechanisms and relationships by which it does so are still largely unknown.

## Syria in Context
The decades preceding the Arab Spring in Syria were marked by repression and
fear. In 1971, Hafez al-Assad overthrew the previous leader and remained in
power for thirty years, presiding during regional contentiousness and
cooperation until his death in 2000. His son, Bashar al-Assad, continued the
regime of authoritarian leadership after assuming his office. The Assad lineage
hails from a minority Islamic sect (the Alawites) in a multi-religious and
multi-ethnic country comprising Christians, Sunni and Shia Muslims, Kurds, and
Palestinian refugees.

@@

Following a wave of democratization efforts across the Middle East and North
Africa, modest Syrian protests began in March 2011. The regime responded harshly
with imprisonments, detentions, censorship, and military operations as the
resistance to Assad's regime spread. The fighting escalated dramatically in
March 2012, after increased unrest in the western city of Homs. International
actors---including the Red Cross/Red Crescent Society, the United Nations
General Assembly, the U.N. Security Council, and individual world
leaders---worked toward cease-fires, negotiations, and concessions without much
success. While many high-level military and diplomatic officials defected,
numerous loyal forces continued the regime's repression, and the Syrian people
turned to social media to coordinate their activism and communicate to the
world. Between March and June 2012, more than 8 million tweets were sent within
Syria, revealing the details of the conflict suppressed by the government.

Syria is a multi-ethnic, multi-lingual, and multi-religious state. Kurdish
people make up between 10-15% of the population, Alawites between 8-15%, and
Christians about 10%. There are also much smaller sects (e.g., Ismailis, Druze),
but Sunni Muslims comprise the vast majority of society. Most people speak
Arabic as their primary language. Most policy reports addressing the current
conflict report on the cohesion of the ruling Alawite military elite, rather
than on the cohesiveness of the opposition movement. However, the Sunni
majority's disaffection with the Syrian government and their lack of political
representation provide sufficient grievance and collective group identification
to make them a cohesive force.

@@

Massive online social cohesion can manifest in real-world social behavior.
Social media platforms enable online social cohesion by offering distributed
information-gathering and real-time information dissemination. Given its
accessibility and proficiency as an information exchange tool, Twitter has
provided social and political activists an opportunity to complement or even
replace real-world social mobilization. For example, during the 2012 Italian
protests for global economic change, the riots in Rome were mirrored by Twitter
activity both on the day of the protests and in the days following them
[@vicari2013public]. Tweeting was determined to be more effective locally in
discussing real-world, real-time events than traditional mainstream media
platforms. As people increasingly rely on social media for local news and
information, we must understand exactly how these micro-blogs connect to and
reflect real-world action. Some argue that these mass movements would not have
been possible without efficient large-scale coordination, both physically on the
ground and virtually on social media platforms.

Along with Facebook, Twitter has been credited as essential to the Arab
Spring---both for protesters and for the government's response to them
[@comninos2011twitter]. The Arab Spring has been deeply examined within the
political science literature [e.g., @lutterbeck2013arab;
@koehler2017political], but to our knowledge, the direct *coupling* between
Syrian Twitter cohesion and international action has not been investigated. In
fact, some have been skeptical about the effect of social media on collective
action at all [@seay2014slacktivism].

This skepticism is summarized in a quote by Lisa @anderson2011demystifying:

@@

Interestingly, these views may help make the case for bridging social media and
social mobilization. Such views suggest the strength of online social ties is
inextricably linked to the *context* and *ambitions* of the actors. Moreover,
research suggests that weak ties still retain considerable political power:
While strong ties tend to bind smaller groups, weak ties tend to bind groups
across distances [@granovetter1977strength].

Social media generates critical information about events, including whether and
how a demonstration will take place, how many people participate, and how those
in power (e.g., state, military) respond. Similarly, these media platforms alert
us to more global shifts within a conflict, such as when fighting moves to a new
city or region, the number of casualties and wounded, and the types of weaponry
used by parties [@zeitzoff2017social]. In monitoring social media channels,
government forces may respond to the perception that citizens are gaining
strength and sympathy. Governments have been shown to be quite tolerant of
dissent so long as it does not metamorphose into collective action
[@king2013censorship].

Citizens communicate with each other with increased transparency due to the
ubiquitous availability of social media such as low-cost and egalitarian
platforms like Twitter. Citizens are able to access these tools via mobile
phones and computers with Internet connectivity, providing a stream of real-time
shared information. Citizens can share the location of events like protests and
military incursions, the degree of state repression, and the number of
casualties. In turn, the government---interested in remaining in
power---monitors social media platforms to ascertain emerging threats
[@comninos2011twitter], such as where new groups are forming, where existing
groups are strengthening, where gatherings are held, and who the leaders are.

@@

\setlength{\leftskip}{0pt}
\end{singlespace}
\noindent With few exceptions, Assad chose to keep social media platforms up and
running. Gohdes [-@gohdes2015pulling] notes that regimes can attempt to thwart
the growth of social protest online and in the streets by shutting down the
Internet. During these times, military repression increases, but government
information about the citizens decreases. As a result, regimes may elect to
maintain Internet connectivity to monitor and surveil their citizens.

@@

The type of collective contentious behavior that manifests in society is a
function of the available political opportunity structures. The political
opportunity structure is “comprised of *(sic)* specific configurations of
resources, institutional arrangements and historical precedents for social
mobilization, which facilitate the development of protest movements in some
instances and constrain them in others” [@kitschelt1986political]. What emerges
as strategies and tactics in contentious politics are a function of
what the system makes possible. This includes the number of people and their
disposition (i.e. nonviolent or violent), as well as their internal structure.
The more cohesive a group is, the greater threat they pose to the establishment.

Cohesion is a familiar concept in the study of civil conflicts, as it pertains
to group affiliation and the ability to overcome collective action problems.
Individuals must act in concert---that is, cohesively---to accomplish the task
of challenging state authority [@olson1965logic;
@eisinger1973conditions; @tilly2004contentious; @francisco1995relationship].
Viewing rebel recruitment during civil conflicts as a function of internal
cohesion, rebel movement success is influenced by the degree of group cohesion,
which itself is a function of geographic distance, ideology, and ethnicity
[@gates2002recruitment]. Within Syria, multiple rebel groups exist, sourcing
fighters locally (from the disaffected Syrian population) and internationally
(from transnational terrorist organizations). Groups compete intensely with one
another for recruitment in the rebel movement and often punish defectors for
nonparticipation [@gates2002recruitment].

@@

The civil war literature has also explored the role of social cohesion. Social
homogeneity increases the likelihood of social conflict, especially when the
society is polarized [@collier2004greed]; this is partly due to nationalist
groups' strong identification with a collective identity [@cederman2007beyond]
and their social solidarity [@sambanis2009s]. These same processes can be
similarly important for rebels [@anyanwu2002economic]. Referencing Gurr's
[-@gurr2000peoples] work on minorities at risk, Regan and Norton
[-@regan2005greed] note that

@@

We contextualize linguistic cohesion with social mobilization and response to
the relative intensity of the conflict. Cohesion is a multi-layered concept
referring to the connectedness of speakers, actions, and societies. It
reflects the continuity of language and expression and shows the
connectivity of participants in a social movement. To address the relationship
between language and discourse and social movements, we adopt an
interdisciplinary approach that incorporates linguistics, psychology, and
political science using nonlinear dynamical systems methodology to disentangle the
relationship between language and action.

@@

The first level of cohesion that we will consider are the words within social
media posts.  In Figure \@ref(fig:cohesion-in-tweets), we see four sequential
tweets of four words each. Each sequential tweet contains some content
included in the previous tweets. The linguistic construct of content word
overlap measures the relationship across units of meaning---here, tweets---where
identical and semantically similar words are repeated [@graesser2004coh]. These
repeated lexical items constitute a theme; on Twitter, this might be a
particular hashtag, a name, a location, or another nominal category. This
localized cohesion is referential; words are repeated and co-referred across
tweets, creating a sense of continuity and cohesiveness.

@@

In estimating cohesion in Twitter posts, tweets that share content words
yield higher cohesion values, even among speakers with opposing
views on the subject matter. Repetition of terms and concepts is captured by
reiteration, which is a type of lexical cohesion in which a subsequent item
repeats a previous item, or refers to it through synonymous language
[@halliday1976cohesion]. In social media, we can conceive of retweets (i.e., reposts
or forwards of previous messages) as a form of reiteration
that increase message cohesion. Since retweets replicate verbatim some part of
another tweet, these tweets would also have higher cohesion. Hashtags function
similarly and are included in our analysis.

@@

The use of similar language can provide insight into *interlocutor cohesion*, or
the degree to which people are talking about similar things. This can include
people who are talking about similar things from radically perspectives, as well
as those who are talking about something from the same perspective. Previous
research from political science has investigated these ideas [@Windsor2020]: The
“Occupy Wall Street” movement has been held up as an example of a movement with
low interlocutor and linguistic cohesion because of its wide range of signs and
slogans, while the “I am a man” protest (during the 1968 sanitation workers'
strike in Memphis, TN) or the “Black Lives Matter” placards (following the
uprising in Ferguson, MO) have high lexical and interlocutor cohesion
[@edgar2018struggle].

Fed by lexical cohesion, interlocutor cohesion ebbs and flows, contextualized by
changing real world events. Interlocutor cohesion may increase preceding an
in-person social mobilization event, alerting a repressive government that
collective mobilization is happening. Interlocutor cohesion may also respond to
exogenous events such as spontaneous, unprovoked militarized conflict.

Given the importance of lexical and interlocutor cohesion for messaging in
political movements, lexical cohesion may be important trace data for
researchers interested in leveraging social media data to analyze social
movements. Similar to large-scale studies of sociocultural movements [sometimes
referred to as *culturomics*; e.g., @michel2011quantitative], political
scientists can track the ways in which conversations on social media coalesce or
diffuse around similar content words. In this way, it matters relatively little
whether the interlocutors have precisely the same views: The point is to
understand whether the society-level conversation is converging on similar
themes.

@@

At the same time, behavioral similarity in certain contexts has been
connected to important psychological and social effects. For example, shared
language is associated with rapport [both in face-to-face and computer-mediated
conversations; @niederhoffer2002linguistic; @riordan2011evidence] and
can even be connected to better performance on shared tasks [as long as
the language is relevant to the task at hand; @fusaroli2012coming]. Shared
movement is associated with greater group bonding and greater willingness
to contribute to the group [@wiltermuth2009synchrony], and shared identity
leads to a greater willingness to adopt the group's goals and
motivations [@walton2012mere].

We expect to see increased linguistic cohesion among people participating in
online discussions about an idea, event, individual, or movement. Given that
Syrian protesters coordinated their activity through Twitter by discussing their
own events (e.g., planned protests, active demonstrations) and their reactions
to events happening around them (e.g., government reactions, international
action or inaction), the cohesion in social media should correlate with the
social cohesion of the group [@zeitzoff2011using].

Social cohesion can have real and wide-ranging impacts on a political movement
and the people in it. A socially cohesive group is inherently more threatening
to the existing regime because it reflects momentum in collective opposition and
the professionalization of alternate leadership. Global insurgency movements
exhibit similar patterns in their social structure [@bohorquez2009common].
Rather than patchwork messages via Twitter, emergent or intentional cohesive
online messaging---using similar language, paralleling in-person messaging
[@Windsor2020]---demonstrates increased organizational structure, inciting a
government response.
